---
title: "Proj2 Untidy Data"
output:
  html_document: default
  pdf_document: default
date: "2023-03-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```
Introduction:  
  
For our week 5 discussion, we were to come forward with datasets exemplifying "untidy" data. Project 2 requires us to take 3 of the peer posted examples from Week 5, tidy the data, and then perform the analysis requested.  
  
This portion will be focusing on NBA Rankings posted by Jacob.  
  
  
 - NBA Rankings  
https://nbarankings.theringer.com/  
Classmate: Jacob Silver  
  
Packages Used:
- tidyverse
  


```{r Import Raw CSVs}
nbaRankings <- read_csv("https://raw.githubusercontent.com/d-ev-craig/DATA607/main/Projects/Project2%20-%20Untidy%20Data/Ringer%20Top%20100%20NBA%20Players%20UNTIDY%20DATA%20-%20Sheet1.csv")

nbaRankings

```


Below we are going to create the columns that we need with the value the column is interested in within the rows. My idea is that as long as each column has at least one record within it that corresponds to its correct variable value, we can filter rows using string recognition. Once I filter those rows, I can create a new column and start cbinding all my columns together.

```{r Creating New Columns}
nbaRankings <- nbaRankings %>% mutate(team=nbaRankings$`2`)%>% mutate(fieldGoalsMade=nbaRankings$`2`)%>% mutate(freeThrowAttempts=nbaRankings$`2`) %>% mutate(assists=nbaRankings$`2`)

nbaRankings <- nbaRankings %>% mutate(position=nbaRankings$`3`)%>% mutate(fieldGoalPercentage=nbaRankings$`3`)%>% mutate(freeThrowPercentage=nbaRankings$`3`) %>% mutate(turnovers=nbaRankings$`3`)

nbaRankings <- nbaRankings %>% mutate(age=nbaRankings$`4`)%>% mutate(threePtrsMade=nbaRankings$`4`)%>% mutate(offRebounds=nbaRankings$`4`) %>% mutate(steals=nbaRankings$`4`)

nbaRankings <- nbaRankings %>% mutate(points=nbaRankings$`5`)%>% mutate(threePtrPercentage=nbaRankings$`5`)%>% mutate(defRebounds=nbaRankings$`5`) %>% mutate(blocks=nbaRankings$`5`)

nbaRankings <- nbaRankings[,-(2:5)]
colnames(nbaRankings)[1] <- 'playerName'
```


```{r}
nbaRankings

#Player Name and Team
playerTeam <- nbaRankings %>% filter(str_detect(nbaRankings$team, '^Team:')) %>% select(playerName,team) #filter rows by detecting strings that contain our data
playerTeam$team <- gsub('Team: ','',playerTeam$team) #clean up the string


#Field Goals Made
fGM <- nbaRankings %>% filter(str_detect(nbaRankings$fieldGoalsMade, '^Field goals made: ')) %>% select(fieldGoalsMade)  #filter rows by detecting strings that contain our data
fGM$fieldGoalsMade <- gsub('Field goals made: ','',fGM$fieldGoalsMade) #clean up the string

playerTeam <- cbind(playerTeam,fGM$fieldGoalsMade) #add our new column
colnames(playerTeam)[3] <- 'fieldGoalsMade' #ensure the name is what we want

# Repeat

#freeThrowAttempts
fTA <- nbaRankings %>% filter(str_detect(nbaRankings$freeThrowAttempts, '^Free throw attempts: ')) %>% select(freeThrowAttempts)
fTA$freeThrowAttempts <- gsub('^Free throw attempts: ','',fTA$freeThrowAttempts)

playerTeam <- cbind(playerTeam,fTA$freeThrowAttempts)
colnames(playerTeam)[4] <- 'freeThrowAttempts'

#assists
fTA <- nbaRankings %>% filter(str_detect(nbaRankings$assists, '^Assists: ')) %>% select(assists)
fTA$assists <- gsub('^Assists: ','',fTA$assists)

playerTeam <- cbind(playerTeam,fTA$assists)
colnames(playerTeam)[5] <- 'assists'

#position
fTA <- nbaRankings %>% filter(str_detect(nbaRankings$position, '^Position: ')) %>% select(position)
fTA$position <- gsub('^Position: ','',fTA$position)

playerTeam <- cbind(playerTeam,fTA$position)
colnames(playerTeam)[6] <- 'position'

#fieldGoalPercentage
fTA <- nbaRankings %>% filter(str_detect(nbaRankings$fieldGoalPercentage, '^Field goal percentage: ')) %>% select(fieldGoalPercentage)
fTA$fieldGoalPercentage <- gsub('^Field goal percentage: ','',fTA$fieldGoalPercentage)

playerTeam <- cbind(playerTeam,fTA$fieldGoalPercentage)
colnames(playerTeam)[7] <- 'fieldGoalPercentage'


#freeThrowPercentage
fTA <- nbaRankings %>% filter(str_detect(nbaRankings$freeThrowPercentage, '^Free throw percentage: ')) %>% select(freeThrowPercentage)
fTA$freeThrowPercentage <- gsub('^Free throw percentage: ','',fTA$freeThrowPercentage)

playerTeam <- cbind(playerTeam,fTA$freeThrowPercentage)
colnames(playerTeam)[8] <- 'freeThrowPercentage'

#turnovers
fTA <- nbaRankings %>% filter(str_detect(nbaRankings$turnovers, '^Turnovers: ')) %>% select(turnovers)
fTA$turnovers <- gsub('^Turnovers: ','',fTA$turnovers)

playerTeam <- cbind(playerTeam,fTA$turnovers)
colnames(playerTeam)[9] <- 'turnovers'

#age
fTA <- nbaRankings %>% filter(str_detect(nbaRankings$age, '^Age: ')) %>% select(age)
fTA$age <- gsub('^Age: ','',fTA$age)

playerTeam <- cbind(playerTeam,fTA$age)
colnames(playerTeam)[10] <- 'age'


#threePtrsMade
fTA <- nbaRankings %>% filter(str_detect(nbaRankings$threePtrsMade, '^3-pointers made: ')) %>% select(threePtrsMade)
fTA$threePtrsMade <- gsub('^3-pointers made: ','',fTA$threePtrsMade)

playerTeam <- cbind(playerTeam,fTA$threePtrsMade)
colnames(playerTeam)[11] <- 'threePtrsMade'

#offRebounds
fTA <- nbaRankings %>% filter(str_detect(nbaRankings$offRebounds, '^Offensive rebounds: ')) %>% select(offRebounds)
fTA$offRebounds <- gsub('^Offensive rebounds: ','',fTA$offRebounds)

playerTeam <- cbind(playerTeam,fTA$offRebounds)
colnames(playerTeam)[12] <- 'offRebounds'

#steals
fTA <- nbaRankings %>% filter(str_detect(nbaRankings$steals, '^Steals: ')) %>% select(steals)
fTA$steals <- gsub('^Steals: ','',fTA$steals)

playerTeam <- cbind(playerTeam,fTA$steals)
colnames(playerTeam)[13] <- 'steals'

#points
fTA <- nbaRankings %>% filter(str_detect(nbaRankings$points, '^Points: ')) %>% select(points)
fTA$points <- gsub('^Points: ','',fTA$points)

playerTeam <- cbind(playerTeam,fTA$points)
colnames(playerTeam)[14] <- 'points'

#threePtrPercentage
fTA <- nbaRankings %>% filter(str_detect(nbaRankings$threePtrPercentage, '^3-pointer percentage: ')) %>% select(threePtrPercentage)
fTA$threePtrPercentage <- gsub('^3-pointer percentage: ','',fTA$threePtrPercentage)

playerTeam <- cbind(playerTeam,fTA$threePtrPercentage)
colnames(playerTeam)[15] <- 'threePtrPercentage'

#defRebounds
fTA <- nbaRankings %>% filter(str_detect(nbaRankings$defRebounds, '^Defensive rebounds: ')) %>% select(defRebounds)
fTA$defRebounds <- gsub('^Defensive rebounds: ','',fTA$defRebounds)

playerTeam <- cbind(playerTeam,fTA$defRebounds)
colnames(playerTeam)[16] <- 'defRebounds'

#blocks
fTA <- nbaRankings %>% filter(str_detect(nbaRankings$blocks, '^Blocks: ')) %>% select(blocks)
fTA$blocks <- gsub('^Blocks: ','',fTA$blocks)

playerTeam <- cbind(playerTeam,fTA$blocks)
colnames(playerTeam)[17] <- 'blocks'

#Final
playerTeam
```

There were a few questions posed for analysis in this post. The following are the ones I chose. "Are the best teams those with the most players in the top 100 regardless of placement? Or is 'one guy' at the top of the mountain enough?"

I'm not sure which teams are the best but what I can do is show the number of times a team appears in a graph. With some contextual knowledge of which teams are the best, a reader could determine rough tendencies.

```{r}
top100Members <- playerTeam %>% arrange(team) %>% count(team) %>% arrange(desc(n))
top100Members
```
So from the chart above we can see BOS, TOR, MIN, LAC, and GSW contain the most players in the top 100. If you go to https://www.teamrankings.com/nba/ and check their ratings we see the following:
  
Boston      : Rank 1  : 7 members  
Golden State: Rank 10 : 5 members  
LA Clippers : Rank 17 : 5 members  
Toronto     : Rank 14 : 5 members  
Minnesota   : Rank 18 : 5 members
  
Some others:  
  
Memphis     : Rank 5  : 4 members
Portland    : Rank 23 : 3 members
Houston     : Rank 29 : 1 member


