---
title: "Week2 Postgres SQL"
author: "Daniel Craig"
date: "2023-02-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(DBI)
```

# Intro  
  
This R Markdown is for connecting to a locally hosted Postgres server using the following packages:
RPostgreSQL
RPostgres  
  
These can be installed with the typical 'install.packages("RPostgreSQL")  
  
These packages will give access to the following important commands:  
 - dbListTables  
 - dbGetQuery  
  
The database we will be connecting to contain data about friends and family ratings of popular/noteworthy 2022 movies and importing that data into a dataframe. All participants were asked to rate movies on a 1 to 5 scale with 5 being the highest rating. Some users had not seen the movie and their data was stored as null values. I allowed NULL values when creating the tables for only the Rating column.
  
## Setup  
  
First, let's establish variables to make connections easy when making our connection call.  
  
I used Postgres to host a SQL Server on my local machine and PSQL to create my tables. I have experience with MS SQL Server and am much more comfortable in that.  
  
I created a Movies table to house names of movies and created a foreign key constraint from my survey results to ensure that the result was from one of the movies selected.  
  
I think next time I will also include another ratings table for valid ratings and potentially a user table to also require those.


```{r Variable Setup}
library(DBI)

db <- "cunydata607"
db_host <- "localhost"
db_port <- "5432"
db_user <- "postgres"
db_pass <- "SuperB!ounce340"

conn <- dbConnect(
  RPostgres::Postgres(),
  dbname = db,
  host = db_host,
  port = db_port,
  user = db_user,
  password = db_pass
)
```

## Connect

Let's check our connection and ensure it's working with dbListTables

```{r Test Con, echo=FALSE}
dbListTables(conn)
```
  
You should have returned a week2_movies tablename.Let's pull our data!

```{r Pull}
dbGetQuery(conn, "SELECT * FROM week2_movies")
```
  
Now that we can pull the info, let's store this in a dataframe in R to play with.  
  
```{r Store}
movies <- dbGetQuery(conn, "SELECT * FROM week2_movies")
movies
```

A note on null values:  
  
When creating the database, I determined that any movies that had not been seen by an individual would be stored as a NULL value and when running any analysis would exclude the values.  
  
A note on ratings:  
  
All participants were asked to rate movies on a 1 to 5 scale with 5 being the highest rating.
