---
title: "DotaNP"
author: "Daniel Craig"
date: '2022-08-07'
output: html_document
---
```{r}
library(ggiraphExtra)
library(knitr)
library(data.table)
library(readr)
library(dplyr)
library(randomForest)
library(caret)
library(ggplot2)
library(gbm)
library(ggsci)
library(tree)
```

8.14 Update:
Finish the Predictions Page
Finish showing filtered data page

Make it prettier

Load In Data
```{r}
#data<-read_csv("_Projects\\Project3\\datasets\\NPBase.csv")

#npData<-read_csv("NPBase.csv")
#npData<- as_tibble(npData)
#npData <- as.data.frame(npData)
#npData$win <- as.character(npData$win)
#recode(npData$win, Yes = "1", No = "0") did not use
#recode(npData$win, "Yes"="TRUE", "No" = "FALSE") did not use

#npData

#npData$win <- npData$win %>% mutate(win = ifelse(win == "No",0,1))


npData <- data.frame(match_id = c(1234,5678,1245),
                      start_time = c(1657339909,1657357190,1657366555),
                      win = c('false','false','true'),
                      hero_id = '53',
                      account_id = c(2345438,259803,438689),
                      leaguename = c('Titus','Destiny','Ultras'),
                      gold_per_min = c(569,549,654),
                      net_worth = c(15770,15317,16985),
                      gold = c(1140,1559,3210),
                      kills = c(4,12,7),
                      tower_damage = c(1245,19823,4599),
                      duration = c(1754,1759,1829),
                      lane = c(1,3,3),
                      lane_role = c(1, 3,3)
                     )


    npData<-read_csv("NPBase.csv")
    npData$gold <- as.numeric(npData$gold)
    dataModel <- data.frame(npData)
    dataModel$win <- as.factor(dataModel$win)
    dataModel$lane <- as.factor(dataModel$lane)
    dataModel$lane_role <- as.factor(dataModel$lane_role)
    dataModel
    
#npData$account_id <- as.factor(npData$account_id)

#npData$match_id <- as.factor(npData$match_id)

# gold      <- "gold"       %in% charvec
#   gold_per_min    <- "gold_per_min"     %in% charvec
#   kills <- "kills"  %in% charvec
#   tower_damage <- "tower_damage" %in% charvec
#   duration <- "duration" %in% charvec
#   lane <- "lane" %in% charvec
#   lane_role <- "lane_role" %in% charvec
#   net_worth <- "net_worth" %in% charvec
    
    
    


```
EDA
```{r Histogram}

varname <- "gold"

varHist <- ggplot(npData, aes(x=gold)) +  
  theme_bw() +                                                                     #Set classic bw plot theme
  geom_histogram(color="black", fill = "#34495E", alpha = 0.8, binwidth = 100) +   #Color options, binwidth set to 100 shares
  labs(x = "Win Or Lose", y = "Count", title = paste0("Counts of ", varname))                                  

varHist
```

```{r Boxplot}
boxPlot <- ggplot(npData, aes(x=win,y=gold), fill = gold) + 
  theme_bw() +                                                              #Set classic bw plot theme
  geom_hline(yintercept = median(npData$gold), size = 0.8) +             #Add line for overall median shares
  geom_point(size = 0.8) +                                                  #Add points
  geom_boxplot(lwd = 0.5, width = 0.5, outlier.size = 0.8, alpha = 0.7) +   #Create boxplot
  xlab("") + ylab("Amount of Gold") +                                             #Label axis
  theme(legend.position = "none") +                                         #Remove legend
  ggtitle(paste0("Wins vs ", varname)) +                                            #Set title
  scale_color_startrek() + scale_fill_startrek()                            #Set color theme

#Display plot
boxPlot
```


```{r NumSum}
# Numerical Summaries

npData

playerMatches <- npData %>% select(win,account_id,match_id)
  
playerMatches

playerWins <- select(playerMatches, -match_id)

#Total Wins by a Player
playerWins <- playerWins %>% group_by(account_id) %>% summarise(TotalWins = sum(win==1))

playerWins <- as.data.table(playerWins)

#Total Matches by a Player

playerMatches <- as.data.table(playerMatches)
playerMatches[, .(rowCount =.N), by=account_id]

playerMatches <- as.data.table(playerMatches)
sumMatches <- playerMatches[, .(totalMatches =.N), by=account_id]

sumMatches <- playerMatches[sumMatches, on =.(account_id =account_id)]

sumMatches

#Player Frame comprising of the player matches and wins
playerFrame <- sumMatches[playerWins, on =.(account_id =account_id)]

playerFrame <- playerFrame %>% mutate(win_perc = TotalWins/totalMatches) %>% arrange(desc(totalMatches))

playerFrame

#League Matches
leagueMatches <- npData %>% select(win,account_id,match_id,leaguename)
leagueMatches <- as.data.table(leagueMatches)
sumLeagueMatches <- leagueMatches[,.(totalLeagueMatches =.N), by=leaguename]

leagueMatches

#League Wins
leagueWins <- leagueMatches %>% group_by(leaguename) %>% summarise(leagueTotalWins = sum(win==1))
leagueWins

leagueWins <- as.data.table(leagueWins)

#League Frame
leagueFrame <- sumLeagueMatches[leagueWins, on =.(leaguename =leaguename)]

leagueFrame <- leagueFrame %>% mutate(win_perc = leagueTotalWins/totalLeagueMatches) %>% arrange(desc(totalLeagueMatches))


#Numerical

avgTable <- table(avgKills=round(mean(npData$kills),digits=2),avgDur=round(mean(npData$duration),digits=2),avgGPM=round(mean(npData$gold_per_min),digits=2),avgNetWorth = round(mean(npData$net_worth),digits=2), avgTowerDmg=round(mean(npData$tower_damage),digits=2))
avgTable

kable(avgTable)


#Categorical

#Create new categorical variables to classify game as high/normal/low GPM, networth, tower damage, kills,duration
table(npData$win,npData$lane,npData$lane_role)

#note that lane role determined by networth


categoryTable <- table(win=npData$win,lane=npData$lane,lanerole=npData$lane_role)
kable(categoryTable)

```

#Modeling Page

## Model Fitting

#Glm Fit
# Change p = user input for proportion of data
# Change var used to user input


```{r Removing Variables per User}

#dataModel <- data.frame(npData)
#dataModel$win <- as.factor(dataModel$win)

if ("account_id" %in% colnames(dataModel)){
 dataModel <- subset(dataModel, select = -account_id) 
}

if ("match_id" %in% colnames(dataModel)){
dataModel <- subset(dataModel, select = -match_id)
}

if ("hero_id" %in% colnames(dataModel)){
dataModel <- subset(dataModel, select = -hero_id)
}

if ("leaguename" %in% colnames(dataModel)){
dataModel <- subset(dataModel, select = -leaguename)
}

if ("start_time" %in% colnames(dataModel)) {
  dataModel <- subset(dataModel, select = -start_time)
}
dataModel

# class(dataModel$win)
# class(dataModel$gold_per_min)
# class(dataModel$net_worth)
# class(dataModel$gold)
# class(dataModel$kills)
# class(dataModel$tower_damage)
# class(dataModel$duration)
# class(dataModel$lane)
# class(dataModel$lane_role)

```
```{r Subsetting Based on User Selected Variables}
keepVars <- c("gold_per_min","kills")
keepVars <- c(keepVars, "win")
dataModel <- dataModel[, names(dataModel) %in% keepVars]
dataModel

```
```{r Partitioning}
dataIndex <- createDataPartition(dataModel$win, p = 0.75, list = FALSE)
class(dataIndex)
dataTrain <- dataModel[dataIndex, ]
dataTest <- dataModel[-dataIndex, ]
```
```{r Logistic Regression}
glmFit <- train(win ~ ., data = dataTrain, 
         method = "glm", 
         family = "binomial",
         preProcess = c("center", "scale"),
         trControl = trainControl(method = "cv", number = 10))

predictions <- predict(glmFit, dataTrain)

devia<-summary(glmFit)$deviance

coef <-summary(glmFit)$coefficients

predictions
devia
coef

confusionMatrix(data = dataTest$win, reference = predict(glmFit, newdata = dataTest))

```

# Classification Tree
# Change . to user input

```{r}
classTreeFit <- tree(win ~ ., data = dataTrain) # The '.' means all variables to be used as explanatory variables
summary(classTreeFit)

class(classTreeFit)
```

### Confusion Matrix of Accuracy
```{r}
df <- data.frame(matrix(rnorm(16), nrow = 4))
colnames(df) <- paste0("Col", 1:4)
rownames(df) <- paste0("Row", 1:4)
datatable(df)
```



###Pruning
```{r Branch Graph}
pruneFit <- cv.tree(classTreeFit, FUN = prune.misclass)
summary(pruneFit)

plot(pruneFit$size, pruneFit$dev, type = "b")
```

```{r Find Best Branch Count}
pruneFit

dfPruneFit <- cbind(size=pruneFit$size,dev=pruneFit$dev)
dfPruneFit <- data.frame(dfPruneFit)
dfPruneFit <- dfPruneFit %>% group_by(size)%>%arrange(size)%>%arrange(dev)
dfPruneFit

bestVal <- dfPruneFit$size[1]
bestVal
```
### Pruned Predictions
```{r}
pruneFitFinal <- prune.misclass(classTreeFit, best = bestVal)
summary(pruneFitFinal)

fullPred <- predict(classTreeFit, dplyr::select(dataTest, -"win"), type = "class")

prunePred <- predict(pruneFitFinal, dplyr::select(dataTest, -"win"), type = "class")

prunePred

```

### Comparison Full Fit
```{r}

cm <- confusionMatrix(fullPred,dataTest$win)

cmFull <- cm$table

cmFull

accFull <- cm$overall[1]

list(cmFull,accFull)
```
### Comparison Pruned Fit
```{r}
pruneTbl <- table(data.frame(prunePred, dataTest[, "win"]))
kable(pruneTbl)
accPrune<-sum(diag(pruneTbl)/sum(pruneTbl))
print(accPrune)
```

## Random Forest
# Change the . on 254 to user input


Here we change our mtry to be ncol(diamondsTrain)/3 since this example was regression
```{r}  
trainRFModel <- train(win ~ ., data = dataTrain,
method = "rf",
trControl = trainControl(method = "repeatedcv", number = 5, repeats = 3),
tuneGrid = data.frame(mtry = sqrt(ncol(dataTrain) - 1)))

trainConMat <- confusionMatrix(trainRFModel, newdata = dataTest)
testConMat <- confusionMatrix(data = dataTest$win, reference = predict(trainRFModel, newdata = dataTest))
trainConMat
testConMat
summary(trainRFModel)

predict(trainRFModel,newdata=dataTest)

```



# Prediction Page Using the GLM

# gold_per_min,net_worth,gold,kills,tower_damage,duration,lane, lane_role
dataModel

predictions <- predict(glmFit, data = c(input$gpm,input$networth,input$gold,input$kills,input$tower_damage,input$duration,input$lane,input$lane_role))

```{r}

dataIndex <- createDataPartition(dataModel$win, p = 0.75, list = FALSE)
dataTrain <- dataModel[dataIndex, ]
dataTest <- dataModel[-dataIndex, ]

glmFit <- train(win ~ ., data = dataTrain, 
         method = "glm", 
         family = "binomial",
         preProcess = c("center", "scale"),
         trControl = trainControl(method = "cv", number = 10))


#Change to be reactive to user data

userData <- data.frame(gold_per_min=600,net_worth=17000,gold=400,kills=11,tower_damage=9000,duration=2500,lane=3,lane_role=3)

userPred<-predict(glmFit, newdata= userData)

userPred


```





