---
title: "Assignment 5 - Untidy Data"
author: "Daniel Craig"
date: "2023-02-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
library(tidyverse)
```
Let's take a first look at what the data looks like before we clean it up.
```{r}
untidyData <- read_csv("https://raw.githubusercontent.com/d-ev-craig/DATA607/main/Week%205%20-%20Untidy%20Data/untidyData.csv")
untidyData
```

So it looks like we have a variable in airline, a variable in arrival (on time vs delayed), a variable in what I'm assuming is a count of flights as the cell values under the destination cities and there's also the variable of destination itself.  
  
At the end I generate a summary count of total flights with delayed arrival for each airline, and print out the tidy dataframe.

```{r}
untidyData[2,1] <- 'Alaska'
untidyData <- untidyData[-3,]
untidyData[4,1] <- 'AM WEST'

pivotLong <- untidyData %>% pivot_longer(cols = c('Los Angeles','Phoenix','San Diego','San Francisco', 'Seattle'), names_to= 'Destination',values_to = 'Count')

pivotLong <- rename(pivotLong, 'airline' = '...1' )
pivotLong <- rename(pivotLong, 'arrivalCategory' = '...2')

pivotLong
# names(pivotLong)[1] <- 'airline'
# names(pivotLong)[2] <- 'arrivalCategory'



pivotLong %>% group_by(airline) %>% filter(arrivalCategory == 'delayed') %>% summarize(sum(Count))

```



For a simple ggplot command, I went ahead and filtered some data to create a new dataframe so that there were no inline computations needed.

```{r}

delays <- pivotLong %>% filter(arrivalCategory == 'delayed')
delays
```


One could make an off the cuff evaluation and use Alaska for all destinations except Seattle.
```{r}
ggplot(delays,aes(x=Destination,y=Count)) + geom_col(aes(fill = airline), position = "dodge")

```


