---
title: "Proj2 Untidy Data Therapists"
output: 
  html_document:
    toc: true
    toc_float: true
  
date: "2023-03-05"
author: "Daniel Craig"

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
library(plyr)
library(dplyr)
```
## Introduction:  
  
  
For our week 5 discussion, we were to come forward with datasets exemplifying "untidy" data. Project 2 requires us to take 3 of the peer posted examples from Week 5, tidy the data, and then perform the analysis requested.  
  
This portion will be focusing on Kory's African American therapists in LA and his provided untidy example.
  
  
Sources:
 - African American Therapists in LA: https://www.psychologytoday.com/us/therapists/ca/los-angeles?category=african-american
Classmate: Kory Martin & his post
  
  
Let's take a look at the data.  
```{r}

therapists <- read_csv("https://raw.githubusercontent.com/d-ev-craig/DATA607/main/Projects/Project2%20-%20Untidy%20Data/therapists.csv")

head(therapists)

therapists
```

### Variables of Interest
The variables that we need to create after taking a look seem to be the following:
- Name  
- Focus
- Credentials  
- Description  
- Phone Number
- Location
- Mode  
- Accepting Clients  
  
We will expand each of the columns one by one to get our values. Grabbing names from the first column will be first.

```{r}
name <- therapists %>% filter(!is.na(X1))
colnames(name)[1] <- 'name'
name
name <- name[,1]
```
### Expanding Second Column  
  
In the second column, we are looking to pull out the following variables
- Specialty
- Description  
- Phone Number  
  
We will approach this by creating sequence vectors that identify the values in the column we are interested in. We then take all those values, turn them into a column, and combine those columns to create our new dataframe. I am choosing this method since pivot_wider is performed within a cell and each variable is at a fixed sequence in the column.

```{r Specialty,Descr}
#coalesce method -not used
#nameExpanded <- name %>% slice(rep(1:n(),each = 4))
#nameExpanded

#therapists2 <- data.frame(nameExpanded,therapists[,-1])
#therapists2

#df %>%
#mutate(A = coalesce(A,B))

#Extract values by sequences
specialSeq <- seq(from=1,to=3200, by = 4)
specialty <- therapists$X2[specialSeq]
therapists2 <- cbind(name,specialty)


#Desc Sequence
descSeq <- seq(from=2,to=3200, by = 4)
desc <- therapists$X2[descSeq]
therapists2 <- data.frame(therapists2,desc)


#phone number Sequence
phoneSeq <- seq(from=3,to=3200, by = 4)
phone <- therapists$X2[phoneSeq]
therapists2 <- data.frame(therapists2,phone)

#The 4th cell in the sequence of column 2 are all NAs
#scraps Sequence
#scrapSeq <- seq(from=4,to=3200, by = 4)
#scrap <- therapists$X2[scrapSeq]
#therapists2 <- data.frame(therapists2,scrap)
head(therapists2)
```
  
### Expanding Third Column  
  
In the third column we want to pull the following variables:  
- Location  
- Online
- Client Acceptance

```{r Location, Online, CLient Acceptance}
#Location Seq
locSeq <- seq(from=1,to=3200, by = 4)
loc <- therapists$X3[locSeq]
therapists2 <- data.frame(therapists2,loc)

#online Seq
onlineSeq <- seq(from=2,to=3200, by = 4)
online <- therapists$X3[onlineSeq]
therapists2 <- data.frame(therapists2,online)

#Client Seq
clientSeq <- seq(from=3,to=3200, by = 4)
client <- therapists$X3[clientSeq]
therapists2 <- data.frame(therapists2,client)

# 4th value in column is all NA
# scrapSeq <- seq(from=4,to=3200, by = 4)
# scrap <- therapists$X3[scrapSeq]
# therapists2 <- data.frame(therapists2,scrap)
# therapists2

```
  
#### NA Handling  
  
  
There are a few columns that contain NA's that I'd prefer to replace with text so that it could easily be readable to an end user.
```{r Normalizing Data & Cleaning Text}
therapists2$specialty[is.na(therapists2$specialty)]<- 'None Listed'
therapists2$phone[is.na(therapists2$phone)]<- 'None Listed'
therapists2$online[is.na(therapists2$online)]<- 'No' # check this column
therapists2$client[is.na(therapists2$client)]<- 'Accepting'

therapists2$online <- str_replace(therapists2$online, '& ','')
```

  
#### Expanding Location Further  
  
  
We will also expand our location column into a state and zip code column.
```{r Pivoting Out City, State, Zip}

therapists2 <- therapists2 %>% separate_wider_delim(loc, delim = ',', names = c('city','var'))

therapists2

therapists2$var <- str_replace(therapists2$var, 'CA ','')

therapists2$var <- trimws(therapists2$var)

therapists2 <- therapists2 %>% rename(c('var' = 'zip'))

state <- rep('CA',800)

therapists2 <- cbind(therapists2,state)

head(therapists2)


#Factorizing
therapists2$zip <- as.factor(therapists2$zip)
therapists2$client <- as.factor(therapists2$client)
therapists2$online <- as.factor(therapists2$online)

```
  
### Analysis  
  
From here its a pretty tidy data set. I did not split name since some of the 'names' are company names and would be best left all as one entity for it to be easily recognizable. I also left unicode inside the descr column since it may be best to leave those if importing the data elsewhere to another platform.  
  
Some analysis that was chosen to be performed:
*Identifying the % of therapist offering online, on-ste*
*Identifying the percent accepting new patients*
*Break down the group based on their zip code*

```{r Analysis}
#Some totals
therapists2 %>% dplyr::group_by(online) %>% dplyr::summarize(n())
therapists2 %>% dplyr::group_by(client) %>% dplyr::summarize(n())
therapists2 %>% dplyr::group_by(zip) %>% dplyr::summarize(total=n()) %>% arrange(desc(total))

#Percentage of Therapists offering online
onlineOffer <- sum(therapists2$online =='Online') #Summing number of therapists offering online
noOnline <- sum(therapists2$online == 'No') #Summing number of therapists not offering online


#percentageNotOnline <- noOnline/800 #800 is the total number of therapists
percentageOnline <- (onlineOffer/800 * 100)
print(paste0('Percentage Offering Online Service is ',percentageOnline, '%'))

newPatients <- sum(therapists2$client =='Accepting') #Summing number of therapists accepting new patients
percentagePatient <- (newPatients/800 * 100)
print(paste0('Percentage Accepting New Patients is ',percentagePatient,'%'))

#Percentage accepting both

both <- sum(therapists2$online == 'Online' & therapists2$client == 'Accepting')
percentageBoth <- (both/800 *100)
print(paste0('Percentage accepting new patients and offering online service is ', percentageBoth,'%'))

# Breaking Groups down based on Zip

g <- ggplot(data=therapists2,aes(x=online, fill=online))
g + geom_bar()


g2 <- ggplot(data=therapists2,aes(x=client, fill=client))
g2 + geom_bar()

#Ordering the table by zip code, then by whether they offer online services, then by whether they're accepting new clients
therapists2 <- therapists2 %>% arrange(zip,online,client)



#Table to help find breakdown by each zip code
therapists2 %>% dplyr::group_by(zip,client,online) %>% dplyr::summarize(total = n())

```
#### Summary  
  
So in summary, of the 800 total therapists:  
- 698 of them offer online service (87.25%)  
- 683 of them are accepting new patients (85.375%), 78 not accepting, 39 have a waitlist  
- Zip code 90025 has the highest total with 80 therapists in the area  