---
title: "Proj2 Untidy Data Therapists"
output: html_document
date: "2023-03-05"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
```
## Introduction:  
#### Daniel Craig
  
For our week 5 discussion, we were to come forward with datasets exemplifying "untidy" data. Project 2 requires us to take 3 of the peer posted examples from Week 5, tidy the data, and then perform the analysis requested.  
  
This portion will be focusing on Kory's African American therapists in LA and his provided untidy example.


Sources:
 - African American Therapists in LA: https://www.psychologytoday.com/us/therapists/ca/los-angeles?category=african-american
Classmate: Kory Martin & his post
 

Let's take a look at the data.
```{r}

therapists <- read_csv("https://raw.githubusercontent.com/d-ev-craig/DATA607/main/Projects/Project2%20-%20Untidy%20Data/therapists.csv")

head(therapists)
```
  
The variables that we need to create after taking a look seem to be the following:
- Name  
- Focus
- Credentials  
- Description  
- Phone Number
- Location
- Mode  
- Accepting Clients  
  
We will expand each of the columns one by one to get our values. Grabbing names from the first column will be first.

```{r}
name <- therapists %>% filter(!is.na(X1))
colnames(name)[1] <- 'name'
name
name <- name[,1]
```
In the second column, we are looking to pull out the following variables
- Specialty
- Description  
- Phone Number  
  
We will approach this by creating sequence vectors that identify the values in the column we are interested in. We then take all those values, turn them into a column, and combine those columns to create our new dataframe. I am choosing this method since pivot_wider is performed within a cell and each variable is at a fixed sequence in the column.

```{r Specialty,Descr}
#coalesce method -not used
#nameExpanded <- name %>% slice(rep(1:n(),each = 4))
#nameExpanded

#therapists2 <- data.frame(nameExpanded,therapists[,-1])
#therapists2

#df %>%
#mutate(A = coalesce(A,B))

#Extract values by sequences
specialSeq <- seq(from=1,to=3200, by = 4)
specialty <- therapists$X2[specialSeq]
therapists2 <- cbind(name,specialty)


#Desc Sequence
descSeq <- seq(from=2,to=3200, by = 4)
desc <- therapists$X2[descSeq]
therapists2 <- data.frame(therapists2,desc)


#phone number Sequence
phoneSeq <- seq(from=3,to=3200, by = 4)
phone <- therapists$X2[phoneSeq]
therapists2 <- data.frame(therapists2,phone)

#The 4th cell in the sequence of column 2 are all NAs
#scraps Sequence
#scrapSeq <- seq(from=4,to=3200, by = 4)
#scrap <- therapists$X2[scrapSeq]
#therapists2 <- data.frame(therapists2,scrap)
head(therapists2)
```

In the third column we want to pull the following variables:  
- Location  
- Online
- Client Acceptance

```{r Location, Online, CLient Acceptance}
#Location Seq
locSeq <- seq(from=1,to=3200, by = 4)
loc <- therapists$X3[locSeq]
therapists2 <- data.frame(therapists2,loc)

#online Seq
onlineSeq <- seq(from=2,to=3200, by = 4)
online <- therapists$X3[onlineSeq]
therapists2 <- data.frame(therapists2,online)

#Client Seq
clientSeq <- seq(from=3,to=3200, by = 4)
client <- therapists$X3[clientSeq]
therapists2 <- data.frame(therapists2,client)

# 4th value in column is all NA
# scrapSeq <- seq(from=4,to=3200, by = 4)
# scrap <- therapists$X3[scrapSeq]
# therapists2 <- data.frame(therapists2,scrap)
# therapists2

```
  
  
There are a few columns that contain NA's that I'd prefer to replace with text so that it could easily be readable to an end user.
```{r Normalizing Data & Cleaning Text}
therapists2$specialty[is.na(therapists2$specialty)]<- 'None Listed'
therapists2$phone[is.na(therapists2$phone)]<- 'None Listed'
therapists2$online[is.na(therapists2$online)]<- 'No' # check this column
therapists2$client[is.na(therapists2$client)]<- 'Accepting'

therapists2$online <- str_replace(therapists2$online, '& ','')
```

  
  
We will also expand our location column into a state and zip code column.
```{r Pivoting Out City, State, Zip}

therapists2 <- therapists2 %>% separate_wider_delim(loc, delim = ',', names = c('city','var'))

therapists2$var <- str_replace(therapists2$var, '\\D{2}','')

therapists2$var <- trimws(therapists2$var)

therapists2 <- therapists2 %>% rename(c('zip' = 'var'))

state <- rep('CA',800)

therapists2 <- cbind(therapists2,state)

head(therapists2)

```

  
From here its a pretty tidy data set. I did not split name since some of the 'names' are company names and would be best left all as one entity for it to be easily recognizable. I also left unicode inside the descr column since it may be best to leave those if importing the data elsewhere to another platform.
